Market-based Mechanisms for Resource Scheduling

### 2013-12-18 paper/paper.tex
worked on intro, added images back

Large-scale computing clusters offer potential for greater efficiency through
pooling of computational resources, but must contend with the problem of
allocating these resources to users with competitive and potentially
incompatible demands. Such conflicts are typically arbitrated by some \emph{fairness}
criterions, such as dominant resource fairness (DRF) \cite{drf}, which benefit from their
simplicity, efficiency, and non-manipulability. Unfortunately, when applied in
realistic settings --- when dealing with inhomogeneous resources, time-varying
demands, requirements for data locality, etc.---these systems cease to be either
efficient or non-manipulable.

We propose using prices as an abstraction for subsystems to communicate their
constraints, and describe a protocol based on auctions for allocating resources
in a cluster. Our protocol has the key advantage that it can continue to meet
efficiency guarantees in realistic environments where preferences are
inhomogeneous across resources and time, and in the presence of adversarial
manipulation. Moreover, we propose prices as a general mechanism for enforcing a
clean abstraction between different components of a system, allowing one
component to make choices intelligently based on constraints elsewhere in the
system even when it was not designed with knowledge of those constraints.

We incorporate these ideas into the Mesos cluster manager, along with microeconomic
``agents'' which allow traditional Mesos clients to interact seamlessly with our
modified implementation. We show through simulations that our market-based mechanism has performance characteristics similar or better compared to existing approaches (in particular, DRF) even when users do not
provide any meaningful price information. Moreover, we show that when accurate
economic data is available our system is able to achieve substantial
performance improvements over DRF. Finally, we discuss a number of examples where this
functionality may be useful even in applications where the end-users of the
system don't have to think about economics at all, and instead prices are
generated algorithmically as a principled way to accommodate constraints that
are currently managed in an ad hoc way.

@@

The abstraction of resource offers allows Mesos to serve frameworks without
understanding the constraints they face. By making offers to those frameworks
which are currently using the least resources, Mesos attempts to ensure that all
frameworks receive a fair share of the cluster. Because of the simplicity of
this protocol, it appears to be immune to manipulation (and indeed it has been
shown to be immune to manipulation in theory).

Unfortunately, while the fairness policy used by Mesos is efficient and
strategy-proof in a simple one-shot game with homogeneous resources, it is
neither efficient nor strategy-proof in this setting. 
%For example, a framework
%using Mesos which prefers to run tasks on machines that already have relevant
%data is forced to engage in strategic behavior: given a resource which it
%\emph{could} use now, should it take the resource or decline it, so that it
%might receive a better offer in the future? Solutions to this problem rely on
%simple heuristics which require tuning based on the parameters of the system;
%there are simple examples where the resulting allocations are not efficient
%despite best-efforts of the clients, and in complex examples there is little to
%believe the results are efficient. 
Moreover, the
current design of Mesos is essentially unsuitable for even mildly hostile
environments. Section \ref{sec:theory} provides detailed theoretical analyses.
%; a framework which can use less than its ``fair share'' of the
%resources at any point in time has no incentive to do so, and indeed has every
%incentive to exploit all resources it can be allocated even if they are being
%used arbitrarily inefficiently (and there is always \emph{something} to do with
%idle resources, if only perform proof-of-work in exchange for small amounts of
%money).

@@

\subsection{Why microeconomic methods?}
The basic paradigm we propose is conceptually simple, though is associated with
many practical difficulties. At each stage where we need to make a decision, we
propose:
\vspace{-5mm}
\begin{enumerate}
\itemsep0em
  \item Estimate the value of different outcomes.
  \item Estimate the outcomes that will result from different decisions.
  \item Compute the value of different decisions.
\end{enumerate}
\vspace{-5mm}  
This paradigm has the virtue of being modular in a number of important respects:
\vspace{-5mm}
\begin{enumerate}
\itemsep0em
  \item The notion of value provides a unifying framework for different
    components to communicate with each other; if I am one component providing a
    service to some client, and I know how valuable different outcomes are to my
    client, I know everything I need to know; my policy does not need to depend
    on anything else about the client.
  \item I can improve and reason about the prediction of outcomes and the
    estimates of value separately. 
  \item It is often possible to compute the value straightforwardly, given the
    values stated  by my clients. For example, if I am scheduling a MapReduce
    instance and know that a 1 second delay costs my client 1 unit, then I can
    estimate the cost of a 1 second delay in one of my tasks simply by
    estimating (or using a crude heuristic for) the probability that this task
    is on the critical path.
\end{enumerate}

\section{Market-based Mechanism Design}

@@

@@

\begin{figure}
%\includegraphics[width=0.5\textwidth]{images/util_high.png}
\caption{Utilization achieved by DRF and our policy on a simulated 25-node CPU-bound cluster.  DRF's average utilization is 98\% average utilization while ours is 99\%.}
\label{img:util_high}
\end{figure}

@@

%\includegraphics[width=0.5\textwidth]{images/util_low.png}
\caption{Behavior of DRF and our policy on a simulated 100-node cluster. Both policies have average utilization of 44\%.}
\label{img:util_low}
\end{figure}

@@

%\includegraphics[width=0.5\textwidth]{images/relative_completion_times.png}
\caption{Relative average job completion times between DRF and auctions for all frameworks.}
\label{img:relative}
\end{figure}

@@

\begin{figure}
%\includegraphics[width=0.5\textwidth]{images/inhomo_times.png}
\caption{The results of simulating Mesos on a cluster of 64 machines, of which 16 have access to a special resource. DRF schedules 30\% of tasks on the optimal machine when frameworks prioritize appropriately, while a simple market-based policy schedules 60\%.}
\label{img:inhomo}
\end{figure}

>>>>

Large-scale computing clusters offer the potential for greater efficiency through
pooling of computational resources, but must contend with the problem of
allocating these resources to users with competitive and potentially
incompatible demands. Such conflicts are typically arbitrated by some \emph{fairness}
criterions, such as dominant resource fairness (DRF) \cite{drf}, which benefit from their
simplicity, efficiency, and non-manipulability. Unfortunately, when applied in
realistic settings --- when dealing with inhomogeneous resources, time-varying
demands or deadlines, requirements for data locality, etc.---typical implementations
of these systems cease to be either
efficient or non-manipulable.

We propose allowing users of the cluster to bid on resources
as a method of expressing their constraints and priorities without exposing
those constraints directly to the scheduler. We imagine that the users
of the cluster will often themselves be programs running on the cluster
and making choices about bidding programmatically, and stress that we are
investigating the suitability of this approach as a communication abstraction
between programs rather than as a user interface to expose to the end-users
of a computing cluster.

Our protocol has the key advantage that it can continue to meet
efficiency guarantees in realistic environments where preferences are
inhomogeneous across resources and time.

We incorporate these ideas into the Mesos cluster manager, along with microeconomic
``agents'' which allow traditional Mesos clients to interact seamlessly with our
modified implementation. We show through simulations that our market-based mechanism 
achieves comparable performance with the existing implementation of Mesos.
We then give an example of modifying the bidding behavior of frameworks
in order to express varying preferences for particular machines in the cluster,
and show that this modification is able to improve the performance of Mesos.

@@

The abstraction of resource offers allows Mesos to serve frameworks without
understanding the constraints they face. By making offers to those frameworks
which are currently using the least resources, Mesos attempts to ensure that all
frameworks receive a fair share of the cluster.

Unfortunately, while the fairness policy used by Mesos is efficient and
strategy-proof in a simple one-shot game with homogeneous resources, it is
neither efficient nor strategy-proof in this more realistic setting. 
%For example, a framework
%using Mesos which prefers to run tasks on machines that already have relevant
%data is forced to engage in strategic behavior: given a resource which it
%\emph{could} use now, should it take the resource or decline it, so that it
%might receive a better offer in the future? Solutions to this problem rely on
%simple heuristics which require tuning based on the parameters of the system;
%there are simple examples where the resulting allocations are not efficient
%despite best-efforts of the clients, and in complex examples there is little to
%believe the results are efficient. 
Moreover, the
current design of Mesos is essentially unsuitable for hostile
environments. Section \ref{sec:theory} provides detailed theoretical analyses.
%; a framework which can use less than its ``fair share'' of the
%resources at any point in time has no incentive to do so, and indeed has every
%incentive to exploit all resources it can be allocated even if they are being
%used arbitrarily inefficiently (and there is always \emph{something} to do with
%idle resources, if only perform proof-of-work in exchange for small amounts of
%money).

@@

\section{Our auction}

@@

Depending on the setting, the budgets of the frameworks may be entirely internal to the cluster
using an idiosyncratic currency,
or the budgets might be shared between several different components of a larger system
(or frameworks may even draw from real budgets of organizations, individuals, or teams,
which have significance outside of the cluster).
The auction we describe is applicable in any of these environments,
though our evaluations all rely exclusively on a closed economy with an idiosyncratic currency.

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{images/util_high.png}
\caption{Utilization achieved by DRF and our policy on a simulated 25-node CPU-bound cluster.  DRF's average utilization is 98\% average utilization while ours is 99\%.}
\label{img:util_high}
\end{figure}

@@

\includegraphics[width=0.5\textwidth]{images/util_low.png}
\caption{Behavior of DRF and our policy on a simulated 100-node cluster. Both policies have average utilization of 44\%.}
\label{img:util_low}
\end{figure}

@@

\includegraphics[width=0.5\textwidth]{images/relative_completion_times.png}
\caption{Relative average job completion times between DRF and auctions for all frameworks.}
\label{img:relative}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{images/inhomo_times.png}
\caption{The results of simulating Mesos on a cluster of 64 machines, of which 16 have access to a special resource. DRF schedules 30\% of tasks on the optimal machine when frameworks prioritize appropriately, while a simple market-based policy schedules 60\%.}
\label{img:inhomo}
\end{figure}
